---
title: "Analyses"
author: "Felix Tran"
date: "December 3, 2018"
output: html_document
---
This file contains code for running OLS regression analyses

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(data.table)
library(tidyverse)
library(ggridges)
library(sp)
library(rgdal)
library(rgeos)
library(ggplot2)
library(broom)
library(gridExtra)
library(spdep)
library(lme4)
```



# Data import

First, have to calculate the average gini of neighboring counties using the code
from Creating_neighbors.Rmd
```{r}
county_df <- readr::read_csv('./Cleaned data/County/final_df.csv')

county_map <- readOGR(dsn = 'cb_2016_us_county_500k',
                   layer = 'cb_2016_us_county_500k',
                   verbose = T)
county_map <- county_map[which(as.integer(as.character(county_map$GEOID)) %in% 
                           county_df$county_code), ]
county_map <- county_map[order(as.integer(as.character(county_map$GEOID))), ]
```

```{r}
neigh_tib <- tibble(
  neighbors = poly2nb(county_map),
  county_code = county_df$county_code
)
```

```{r}
calc_neigh_gini <- function(neigh_vector) {
  total_gini <- 0
  num_neighbors <- length(neigh_vector)
  
  for (i in 1:num_neighbors) {
    total_gini <- total_gini + 
      county_df$gini[neigh_vector[i]]
  }
  
  neigh_gini <- round(total_gini / num_neighbors, digits = 4)
  neigh_gini
}

county_df <- county_df %>% 
  left_join(., neigh_tib, by = "county_code") %>% 
  mutate(neigh_gini = purrr::map(neighbors, calc_neigh_gini))

for (i in 1:nrow(county_df)) {
  if (length(unlist(county_df$neigh_gini[i])) > 0)
    county_df$neigh_gini[i] <- unlist(county_df$neigh_gini[i])
  else
    county_df$neigh_gini[i] <- 0
}

county_df$neigh_gini <- unlist(county_df$neigh_gini)
```

Read in state-level data for multilevel model
```{r}
state_df <- readr::read_csv('./Cleaned data/State/state_final_df.csv')

get_state_gini <- function(county_id) {
  state_df$gini[which(state_df$state_code == as.integer(county_id / 1000))]
}

county_df <- county_df %>% 
  mutate(state_gini = purrr::map(county_code, get_state_gini))

county_df$state_gini <- unlist(county_df$state_gini)
```


Only look at counties with usable suicide rates
```{r}
final_df <- county_df %>% 
  filter(usable == T) %>% 
  mutate(gini = gini * 100) %>%
  # mutate(neigh_gini = neigh_gini * 100) %>%
  # mutate(state_gini = state_gini * 100) %>%
  mutate(median_income = round(median_income / 10000, digits = 3)) %>% 
  mutate(urban_rural_code = as.integer((urban_rural_code - 1) / 2)) %>% 
  mutate(urban_rural_code = as.factor(urban_rural_code)) 
# %>% 
#   select(gini, age_adjusted_rate, male_prop, white_prop, geriatric_prop,
#          college_prop, poverty_prop, urban_rural_code, everything())
```



# OLS Regression

### Crude model - exposure and outcome only
Starting with a simple ordinary least squares (OLS) linear regression modeling
the relation between age-adjusted suicide rates and income inequality as 
measured by the gini coefficient among US counties
```{r}
crude_model <- lm(data = final_df, crude_rate ~ gini) %>% 
  broom::tidy() %>% 
  mutate(estimate = round(estimate, digits = 2),
         std.error = round(std.error, digits = 2),
         statistic = round(statistic, digits = 2),
         p.value = round(p.value, digits = 3)) %>% 
  mutate(lower_ci = round(estimate - 1.96*std.error, digits = 2),
         upper_ci = round(estimate + 1.96*std.error, digits = 2),
         ci = str_c(lower_ci, upper_ci, sep = ", "))

crude_model
```


##### Scatterplot of main exposure and outcome
```{r}
final_df %>% 
  ggplot(aes(x = gini, y = crude_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = lm, alpha = 0.75)
```



### Model with main exposure and average gini coefficient of neighboring counties
```{r}
model_two <- lm(data = final_df, crude_rate ~ gini + neigh_gini) %>% 
  broom::tidy() %>% 
  mutate(estimate = round(estimate, digits = 2),
         std.error = round(std.error, digits = 2),
         statistic = round(statistic, digits = 2),
         p.value = round(p.value, digits = 3)) %>% 
  mutate(lower_ci = round(estimate - 1.96*std.error, digits = 2),
         upper_ci = round(estimate + 1.96*std.error, digits = 2),
         ci = str_c(lower_ci, upper_ci, sep = ", "))

model_two
```



### Model with main exposure, average gini of neighbors, and state gini
```{r}
model_three <- lme(data = final_df, crude_rate ~ gini + neigh_gini +
                      state_gini, random = ~1|state) 

summary(model_three)
```


Testing multicollinearity
```{r}
gini_df <- final_df %>% 
  dplyr::select(gini, neigh_gini, state_gini)

gini_multicol_matrix <- round(stats::cor(gini_df), digits = 3)
gini_multicol_matrix
```



### Model with main exposure and covariates
```{r}
model_four <- lm(data = final_df, crude_rate ~ gini + 
                   median_age +
                   below_hs_prop + college_prop + 
                   male_prop + 
                   black_prop + native_prop + asian_prop + hispanic_prop + 
                   urban_rural_code +
                   unemployed_prop +
                   median_income) %>% 
  broom::tidy() %>% 
  mutate(estimate = round(estimate, digits = 2),
         std.error = round(std.error, digits = 2),
         statistic = round(statistic, digits = 2),
         p.value = round(p.value, digits = 3)) %>% 
  mutate(lower_ci = round(estimate - 1.96*std.error, digits = 2),
         upper_ci = round(estimate + 1.96*std.error, digits = 2),
         ci = str_c(lower_ci, upper_ci, sep = ", "))

model_four
```

# Testing significance for categorical variable
```{r}
l_full <- logLik(lm(data = final_df, crude_rate ~ gini + 
                   median_age +
                   below_hs_prop + college_prop + 
                   male_prop + 
                   black_prop + native_prop + asian_prop + hispanic_prop + 
                   # urban_rural_code +
                   median_income))

l_restricted <- logLik(lm(data = final_df, crude_rate ~ gini + 
                   median_age +
                   below_hs_prop + college_prop + 
                   male_prop + 
                   black_prop + native_prop + asian_prop + hispanic_prop + 
                   unemployed_prop +
                   median_income))

1 - pchisq(2*(l_full - l_restricted), df = 2)
```




# Multilevel model - county and state gini
```{r}
summary(lmer(crude_rate ~ gini + (1 | state), data = final_df, REML = F))
```


# Multilevel model with confounders
```{r}
# summary(lmer(crude_rate ~ gini + state_gini + (gini | state) + median_age + 
#                below_hs_prop + college_prop + male_prop + black_prop + 
#                native_prop + asian_prop + hispanic_prop + urban_rural_code +
#                median_income, 
#              data = final_df, REML = F))

summary(lmer(crude_rate ~ gini + (1 | state) +
               median_age +
               below_hs_prop + college_prop +
               male_prop +
               black_prop + native_prop + asian_prop + hispanic_prop +
               urban_rural_code +
               unemployed_prop +
               median_income,
             data = final_df, REML = T))
```
