---
title: "Data_cleaning"
author: "Felix Tran"
date: "October 14, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
```

This file documents all the data cleaning and processing for my thesis. Data
from 4 different sources were used. The data are categorized as 1 of 4 kinds -
exposure data, outcome data, covariate data, or mediator data.

# Exposure data - ACS 2012-2016 5-year estimates
County-level income inequality was measured using a Gini coefficient and was
retrieved from the American Community Survey 2012-2016 5-year estimates.

These are the steps I took to clean and process this data:

1. Removed an unneeded ID column

2. Renamed variables for ease of use

3. Created a new variable to separate county name and state
```{r exposure data}
acs_exp_df <- read_csv('./Raw data/County/ACS 2012-2016 Gini.csv', skip = 1) %>% 
  janitor::clean_names() %>% 
  select(-id, 
         -margin_of_error_gini_index,
         county_code = id2, 
         county = geography, 
         gini = estimate_gini_index) %>% 
  separate(county, into = c('county', 'state'), sep = ',') %>% 
  mutate(county_code = as.integer(county_code))
```




# Outcome data - CDC WONDER
Suicide rate by county data were retrieved from the CDC WONDER Compressed
Mortality data and exported as a .txt file. Details about the options used when
exporting the data are documented in "CDC wonder query criteria.docx".

These are the steps I took to clean and process the data:

1. Wrote a function to replace suppressed or unreliable rates with NA's

2. The .txt file contained notes about the data at the end of the file. 
These notes were deleted from the dataset, but they can be viewed in 
"CDC wonder notes.docx".

3. I created variables to note if the cruderate for a particular county
was suppressed or unreliable and therefore unusable, or not.

4. Converted the usable crude and age-adjusted rates into numeric
```{r outcome data}
clean_rates <- function(x) {
  unusable <- which(str_detect(x, "Suppressed"))
  unrel <- which(str_detect(x, "Unreliable"))
  x[unusable] <- NA
  
  for (i in 1:length(unrel)) {
    x[unrel[i]] <- str_sub(x[unrel[i]], end = 5)
  }
  x
}

cdc_df <- readr::read_tsv('./Raw data/County/Compressed Mortality, 1999-2016.txt') %>% 
  janitor::clean_names() %>% 
  drop_na(county) %>% 
  dplyr::select(-notes) %>% 
  mutate(suppressed = if_else(deaths == "Suppressed", T, F),
         unreliable = 
           if_else(str_detect(crude_rate, "Unreliable"), T, F),
         usable = !(suppressed | unreliable)) %>% 
  dplyr::select(county_code, 
         age_adjusted_rate, 
         crude_rate,
         suppressed,
         unreliable,
         usable) %>% 
  mutate(age_adjusted_rate = clean_rates(age_adjusted_rate),
         crude_rate = clean_rates(crude_rate)) %>% 
  mutate(county_code = as.integer(county_code),
         age_adjusted_rate = as.numeric(age_adjusted_rate),
         crude_rate = as.numeric(crude_rate))

usable_counties <- length(which(cdc_df$usable))
suppressed_counties <- length(which(cdc_df$suppressed))
unreliable_counties <- length(which(cdc_df$unreliable))
```

After this initial process I am left with age-adjusted suicide rates for 
`r usable_counties` counties.




# Covariates
Covariates for this analysis are: race/ethnicity, gender, county 
urban/ruralness, educational attainment, age, absolute income level, and 
poverty.

Studies and vital stats have shown that age-adjusted suicide rates differ based
on these variables and thus should be adjusted for in analyses.

### Urban/rural classification
Urban/rural classification is determined by the NCHS. The NCHS most recently
classified counties in 2013.

These are the steps I took to clean the data:

1. Keep only the county code and urban/rural classification

2. Changed variable names for ease of use
```{r covariate urban/rural}
nchs_df <- readxl::read_excel('./Raw data/County/NCHSURCodes2013.xlsx') %>% 
  janitor::clean_names() %>%
  select(county_code = fips_code,
         urban_rural_code = x2013_code) %>% 
  mutate(county_code = as.integer(county_code))
```


### Economic characteristics - Median income, poverty, unemployment, occupation

Data on county economic characteristics were collected from the ACS 2012-2016 
5 year data. Past studies have suggested that low/poor economic conditions
lead to worse health outcomes. Observed associations between income inequality
and suicide may be attributable to poor economic conditions and not necessarily
inequality within a county i.e. suicides within a county would decrease if 
absolute/overall economic conditions improved even if gaps between social/income
groups increase.

The variables selected as economic measures are: median income, poverty, 
unemployment among civilians in the labor force ,and proportion of employed 
individuals working in "blue-collar" occupations (natural resources, 
construction, maintenance, production, transportation, and material moving).

These are the steps I took to clean the data:

1. Deleted the 1st row of the dataset which contains unneeded notes

2. Only keep information for county code, county name, median income, poverty, 
unemployment, and blue-collar occupations
```{r covariate absolute income/poverty}
econ_df <- read_csv('./Raw data/County/ACS 2012-2016 Economics.csv')
econ_df <- econ_df[-1, ]
econ_df <- econ_df %>% 
  janitor::clean_names() %>%
  select(county_code = geo_id2,
         median_income = hc01_vc86,
         poverty_prop = hc03_vc171,
         unemployed_prop = hc03_vc12,
         hc03_vc44, 
         hc03_vc45) %>% 
  mutate(county_code = as.integer(county_code),
         blue_collar_prop = as.numeric(hc03_vc44) + as.numeric(hc03_vc45)) %>% 
  select(-hc03_vc44, -hc03_vc45)
```


### Educational attainment

Data on educational attainment were obtained from the ACS 2012-2016 5 year data.
Differences in suicide rates by educational attainment have been observed in 
the literature. Educational attainment in a county could also be used as a
measure of SES.

Data were obtained on the proportion of people in a county with less than a
high school education, high school/GED education, some college/associate degree,
and college degree or above among people 25 years and older.

These are the steps I took to clean the data:

1. Only kept variables for the county code, % of people with no high school 
diploma/equivalent, % of people with high school/GED education, % of people
with some college/associate degree, and % of people with college degree or 
greater among people 25 years and older.

2. Remove an unneeded line of notes from the dataset
```{r covariate education}
edu_df <- read_csv('./Raw data/County/ACS 2012-2016 Education.csv') %>% 
  janitor::clean_names() %>% 
  select(county_code = geo_id2,
         hc02_est_vc09,
         hc02_est_vc10,
         hc02_est_vc11,
         hc02_est_vc12,
         hc02_est_vc13,
         hc02_est_vc14,
         hc02_est_vc15) %>% 
  mutate(county_code = as.integer(county_code),
         below_hs_prop = as.numeric(hc02_est_vc09) + as.numeric(hc02_est_vc10),
         hs_prop = as.numeric(hc02_est_vc11) + as.numeric(hc02_est_vc12),
         college_prop = as.numeric(hc02_est_vc13) + as.numeric(hc02_est_vc14) +
           as.numeric(hc02_est_vc15)) %>% 
  select(-c(hc02_est_vc09, hc02_est_vc10, hc02_est_vc11, hc02_est_vc12, 
            hc02_est_vc13, hc02_est_vc14, hc02_est_vc15))
edu_df <- edu_df[-1, ]
```


### Demographics - Race/ethnicity, gender, and age
Race/ethnicity, gender, and age were retrieved from the ACS 2012-2016 5 year 
estimate data. Studies and vital stats have indicated that non-Hispanic whites 
have higher age-adjusted rates of suicide compared to other racial/ethnic
groups, males have greater age-adjusted rates of suicide compared to females,
and the elderly (65 years and older) have greater rates of suicide compared to
other age groups. To capture any variations, I extracted data on other race
and ethnic group and age groups

These are the steps I took to clean the data:

1. Only kept variables for county code, % of total population that identified
as white, % of total population that is male, % of total population that is 65
years and older, and the margins of error for these estimates

2. Removed a row of unneeded notes from the data
```{r covariate gender/race/age}
demographic_df <- read_csv('./Raw data/County/ACS 2012-2016 Demographics.csv') %>% 
  janitor::clean_names() %>% 
  select(county_code = geo_id2,
         male_prop = hc03_vc04,
         white_prop = hc03_vc94,
         black_prop = hc03_vc95,
         native_prop = hc03_vc96,
         asian_prop = hc03_vc97,
         hispanic_prop = hc03_vc88,
         median_age = hc01_vc23,
         geriatric_prop = hc03_vc29,
         hc03_vc11, hc03_vc12,
         hc03_vc13, hc03_vc14, 
         hc03_vc15, hc03_vc16, hc03_vc17) %>% 
  mutate(county_code = as.integer(county_code),
         age_15_24_prop = as.numeric(hc03_vc11) + as.numeric(hc03_vc12),
         age_25_44_prop = as.numeric(hc03_vc13) + as.numeric(hc03_vc14),
         age_45_64_prop = as.numeric(hc03_vc15) + as.numeric(hc03_vc16) +
           as.numeric(hc03_vc17)) %>% 
  select(-c(hc03_vc11, hc03_vc12, hc03_vc13, hc03_vc14, hc03_vc15,
            hc03_vc16, hc03_vc17))
demographic_df <- demographic_df[-1, ]
```




# Mediator - Social capital
Social capital was measured by the method described in Rupasingha et al (2006).
```{r mediator data}
socialcap_df <- readxl::read_excel('./Raw data/County/Social capital 2014.xlsx') %>% 
  janitor::clean_names() %>% 
  mutate(county_code = as.integer(fips)) %>% 
  select(-fips, -county_name)
```




# Combining datasets
This section combines all the datasets into 1 cohesive dataset containing all
exposure, outcome, covariate, and mediator information.

The Federal Information Processing Standard Publication 6-4 is a 5-digit code 
used to identify counties/county equivalents in the US. Datasets will be 
merged by this county code.
```{r merge datasets}
final_df <- left_join(acs_exp_df, cdc_df, by = "county_code") %>% 
  left_join(., econ_df, by = "county_code") %>% 
  left_join(., nchs_df, by = "county_code") %>% 
  left_join(., edu_df, by = "county_code") %>% 
  left_join(., socialcap_df, by = "county_code") %>% 
  left_join(., demographic_df, by = "county_code")
```

I then saved the final dataset as a csv file for future analyses.
```{r save final dataset}
write_csv(final_df, path = "./Cleaned data/County/final_df.csv")
```


