---
title: "Descriptive_statistics"
author: "Felix Tran"
date: "December 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
```

This file contains code for obtaining descriptive statistics for the variables
used in analyses

# Function to calculate descriptive statistics

The function calculates the mean, standard deviation, min, and max values for 
a given variable.
```{r}
descriptive_stats <- function(var_col) {
  stats_tibble = tibble(
    mean = mean(var_col),
    std_dev = sd(var_col),
    min = min(var_col),
    max = max(var_col),
    range = str_c(min, max, sep = " - ")
  )
  
  stats_tibble
}
```

# Data import

```{r}
final_df <- readr::read_csv('./Cleaned data/final_df.csv') %>% 
  filter(usable == T) %>% 
  select(urban_rural_code,gini, age_adjusted_rate, poverty_prop, college_prop, 
         sk2014, male_prop, white_prop, geriatric_prop)
```

# Calculating descriptive stats

### Urban-rural code

Proportions for urban-rural code were calculated separately because the variable
is coded as a categorical variable. 

There are only 68 counties (4% of analytic sample) with rural code 1 (Most 
rural) included in analyses. The codes will be aggregated into 3 groups - 
Urban = 2 (5, 6), in-between = 1 (3,4), and rural = 0 (1,2). 
```{r}
final_df %>% 
  select(urban_rural_code) %>% 
  mutate(urban_rural_recoded = as.integer((urban_rural_code - 1) / 2)) %>% 
  group_by(urban_rural_recoded) %>% 
  summarize(counts = n()) %>% 
  mutate(proportion = counts/sum(counts))
```

### Remaining variables

The remaining variables are continuous and therefore can be calculated in the
same fashion.
```{r}
raw_calc <- final_df %>% 
  select(-urban_rural_code) %>% 
  summarize_all(funs(mean, sd, min, max))
```

```{r}
select_col <- function(name) {
  col = raw_calc %>% 
    select(starts_with(name))
  
  col
}

df <- tibble(
  var = purrr::map(c('gini', 'age', 'sk', 'white', 'male', 'geriatric',
                     'college'), select_col)
)


```

