---
title: "Exploratory_analysis_and_visualization"
author: "Felix Tran"
date: "October 27, 2018"
output:
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(sp)
library(rgdal)
library(rgeos)
library(ggplot2)
library(broom)
library(RColorBrewer)
library(gridExtra)
library(readr)
```

This file contains exploratory and visualization analyses done before any
formal statistical testing. 

# Loading the dataset

Loads the final dataset created from the data_cleaning file.
```{r loading dataset, include = F}
final_df <- read_csv('./Cleaned data/County/final_df.csv')
state_df <- read_csv('./Cleaned data/State/state_final_df.csv')
merged_df <- read_csv('./Cleaned data/merged_df.csv')
```

Get counts of usable, non-usable counties.
```{r, echo = F}
usable_num <- length(which(final_df$usable == T))
not_usable_num <- length(which(final_df$usable == F))
suppressed_num <- length(which(final_df$suppressed == T))
unreliable_num <- length(which(final_df$unreliable == T))
```

# Exposure data - Visualization

Looking at the distribution of gini coefficients across all counties
```{r exposure visualization, echo = F}
gini_df <- final_df %>%
  filter(usable == T) 

set.seed(23)
normal_df <- tibble(
  observation = c(1:usable_num),
  gini = rnorm(usable_num, mean(gini_df$gini, na.rm = T), 
               sd(gini_df$gini, na.rm = T)))

ggplot() +
  geom_histogram(data = gini_df, aes(x = gini), binwidth = 0.001, alpha = 0.6) +
  geom_density(data = normal_df, aes(x = gini, stat(density)),
               color = 'red', size = 1) +
  theme_minimal() +
  labs(title = "Histogram of usable gini coefficients
against a hypothetical normal curve")

```


# Outcome data - Descriptive/summary statistics and visualizations
1 assumption of linear regression is that the residuals/error terms are normally
distributed. This assumption would be violated if the distribution of
age-adjusted suicide rates is not normally distributed. Therefore checking the
distribution of the included outcome data is important.


### Summary and descriptive statistics 
Out of the 3142 counties in the US, `r usable_num` have age-adjusted suicide 
rates for the timeframe 2012-2016 that are usable. The remaining 
`r not_usable_num` are either suppressed (`r suppressed_num`) or unreliable 
(`r unreliable_num`).
```{r outcome descriptive, echo = F}
# Summary statistics for the usable age-adjusted suicide rates
summary(final_df$age_adjusted_rate)
summary_stats <- final_df %>% 
  summarize(mean = mean(age_adjusted_rate, na.rm = T),
            sd = sd(age_adjusted_rate, na.rm = T))
summary_stats


# See the proportion of suppressed/unreliable rates by state
excluded_df <- final_df %>%
  group_by(state) %>%
  summarize(n_counties = n(),
            n_excluded = sum(usable == F, na.rm = T),
            prop_missing = round(n_excluded / n_counties, digits = 2)) %>%
  arrange(prop_missing)
options(tibble.print_max = Inf)
print(excluded_df)
# write_csv(excluded_df, path = './Cleaned data/County/excluded_df.csv')
```

### Plotting
```{r excluded outcomes plotting, echo = F, fig.height = 8}
# See proportions of usable/non-usable county rates by state
ordering <- excluded_df$state[order(excluded_df$prop_missing)]

excluded_df %>% 
  mutate(n_included = n_counties - n_excluded) %>% 
  gather(key = 'County', value = 'num', c(n_included, n_excluded)) %>%
  mutate(state = parse_factor(state, ordering)) %>% 
  ggplot(aes(x = state, y = num, fill = County), position = 'stack') +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8, color = 'black')) +
  labs(title = "Number of counties included/excluded
       from analysis")

excluded_df %>% 
  mutate(prop_included = 1 - prop_missing) %>% 
  rename(prop_excluded = prop_missing) %>% 
  gather(key = 'proportion', value = 'rate', c(prop_included, prop_excluded)) %>%
  mutate(state = parse_factor(state, ordering)) %>%
  ggplot(aes(x = state, y = rate, fill = proportion), position = 'stack') +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8, color = "black")) +
  labs(title = "Proportion of counties included/excluded 
       from analysis",
       y = "Proportion (%)",
       x = "State")


```

```{r outcome plotting, echo = F}
# Sample 1684 observations from a hypothetical normal distribution with the 
#  same mean and standard deviation as our actual sample of usable suicide rates
set.seed(23)
normal_df <- tibble(
  observation = c(1:usable_num),
  rates = rnorm(usable_num, summary_stats$mean, summary_stats$sd)) 


# Histogram and normal curve
ggplot() +
  geom_histogram(data = final_df, aes(x = age_adjusted_rate),
                 alpha = 0.6, binwidth = 1) +
  geom_density(data = normal_df, aes(x = rates, stat(count)),
               color = 'red', size = 1) +
  theme_minimal() +
  labs(title = "Histogram of usable age-adjusted suicide rates 
against a hypothetical normal curve")
```

Histogram of age-unadjusted suicide rates
```{r}
summary_stats <- final_df %>% 
  summarize(mean = mean(crude_rate, na.rm = T),
            sd = sd(crude_rate, na.rm = T))

set.seed(23)
normal_df <- tibble(
  observation = c(1:usable_num),
  rates = rnorm(usable_num, summary_stats$mean, summary_stats$sd)) 


# Histogram and normal curve
ggplot() +
  geom_histogram(data = final_df, aes(x = crude_rate),
                 alpha = 0.6, binwidth = 1) +
  geom_density(data = normal_df, aes(x = rates, stat(count)),
               color = 'red', size = 1) +
  theme_minimal() +
  labs(title = "Histogram of crude suicide rates against a hypothetical normal curve")
```



# Plotting covariates and outcome

Creating plots of covariates by outcome to see if any covariates should be
transformed or recoded for the linear regression model
```{r}
final_df <- final_df %>% 
  filter(usable == T) %>% 
  select(gini, age_adjusted_rate, male_prop, white_prop, geriatric_prop,
         college_prop, poverty_prop, urban_rural_code)

# Sex
ggplot(data = final_df, aes(x = male_prop, y = age_adjusted_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth(alpha = 0.75)

# Race
ggplot(data = final_df, aes(x = white_prop, y = age_adjusted_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth(alpha = 0.75)

# Education
ggplot(data = final_df, aes(x = college_prop, y = age_adjusted_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth(alpha = 0.75)

# Age
ggplot(data = final_df, aes(x = geriatric_prop, y = age_adjusted_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth(alpha = 0.75)

# Poverty
ggplot(data = final_df, aes(x = poverty_prop, y = age_adjusted_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth(alpha = 0.75)

# Urban-rural code
final_df %>% 
  mutate(urban_rural_code = as.factor(urban_rural_code)) %>% 
  ggplot(aes(x = urban_rural_code, y = age_adjusted_rate)) +
  geom_boxplot()

# Urba-rural (collapsed into 3 categories)
final_df %>% 
  mutate(urban_rural_recoded = as.integer((urban_rural_code - 1) / 2)) %>% 
  mutate(urban_rural_recoded = as.factor(urban_rural_recoded)) %>%
  ggplot(aes(x = urban_rural_recoded, y = age_adjusted_rate)) +
  geom_boxplot()
```

The graphs of the crude relationships between each covariate and suicide rates
are overall in line with my expectations. For all the covarites, there are a
handful of outliers, but I feel the association as a whole won't be changed
by these outliers. Asides from collapsing urban_rural_code into 3 categories
since 1 category has very few observations compared to the rest, I won't
transform the covariates.


### State-level 
```{r}
state_df %>% 
  ggplot(aes(x = ffl_per_100000, y = crude_rate)) +
  geom_point() +
  geom_smooth(method = lm)

state_df %>% 
  ggplot(aes(x = ffl_per_100000, y = age_adjusted_rate)) +
  geom_point() +
  geom_smooth(method = lm)
```



