---
title: "GWR_analyses"
author: "Felix Tran"
date: "February 13, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

This file contains code for running geographically weighted regression (GWR) and
plotting the results of the GWR analysis.

# Load R packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(data.table)
library(broom)
library(RColorBrewer)
library(gridExtra)
library(GWmodel)
library(mapproj)
library(maptools)
library(rgdal)
```



# Read in data (Datasets, maps)
```{r}
# Data
county_df <- readr::read_csv('./Cleaned data/County/final_df.csv') %>% 
  mutate(gini = gini * 100) %>% 
  mutate(median_income = round(median_income / 10000, digits = 3)) %>% 
  mutate(urban_rural_code = as.integer((urban_rural_code - 1) / 2)) %>% 
  mutate(urban_rural_code = as.factor(urban_rural_code)) %>% 
  arrange(county_code)


# Maps (County-level and national)
county_map <- readOGR(dsn = 'cb_2016_us_county_500k',
                   layer = 'cb_2016_us_county_500k',
                   verbose = T)
county_map <- county_map[which(as.integer(as.character(county_map$GEOID)) %in% 
                           county_df$county_code), ]
county_map <- county_map[order(as.integer(as.character(county_map$GEOID))), ]

county_map@data$GEOID <- as.numeric(as.character(county_map@data$GEOID))

county_map_merged <- merge(county_map, county_df, 
                        by.x = 'GEOID', by.y = 'county_code')

county_map_merged$id <- sapply(slot(county_map_merged, "polygons"),
                           function(x) slot(x, "ID"))

county_map_merged <- county_map_merged[which(county_map_merged@data$usable == T), ]


nation_map <- readOGR(dsn = 'cb_2016_us_state_500k',
                   layer = 'cb_2016_us_state_500k',
                   verbose = T)
nation_map <- nation_map[-which(nation_map$NAME %in% 
                            c('American Samoa',
                              'Commonwealth of the Northern Mariana Islands',
                              'Guam', 
                              'United States Virgin Islands',
                              'Puerto Rico')),]
nation_map <- nation_map[order(nation_map$NAME), ]
```



# Perform GWR

Testing different bw values, 102 was the smallest value in which all local
regression equations were solvable.

After performing GWR, I looked at the descriptive stats of the regression
coefficients and prepared files for plotting in ggplot.
```{r}
# GWR
gwr_result <- gwr.basic(crude_rate ~ gini + median_age + below_hs_prop + 
                          college_prop + male_prop + black_prop + native_prop + 
                          asian_prop + hispanic_prop + urban_rural_code + 
                          unemployed_prop + median_income, 
                        data = county_map_merged, adaptive = T, 
                        bw = 102, kernel = 'boxcar')
gwr_result

gwr_df <- gwr_result$SDF

mean(gwr_df@data$gini)
mean(gwr_df@data$gini_SE)


gwr_df$id <- sapply(slot(gwr_df,  "polygons"), function(x) slot(x, "ID"))

gwr_tidied <- tidy(gwr_df)
gwr_tidied <- merge(gwr_tidied, gwr_df@data, by = "id", all.x = T)


# Prepare data from excluded counties for plotting in ggplot
county_map@data$GEOID <- as.numeric(as.character(county_map@data$GEOID))

county_map_merged <- merge(county_map, county_df, 
                        by.x = 'GEOID', by.y = 'county_code')

county_map_merged$id <- sapply(slot(county_map_merged,  "polygons"),
                           function(x) slot(x, "ID"))

county_map_merged <- county_map_merged[which(county_map_merged@data$usable == F), ]

county_tidied <- tidy(county_map_merged)

county_tidied <- merge(county_tidied, county_map_merged@data, by = "id",
                        all.x = T) %>% 
  dplyr::select(id, long, lat, order, hole, piece, group)
```



# Save shapefile of GWR results

I saved the shapefile of GWR results. I then imported the shapefile in QGIS to 
create a new shapefile of the statistically significant clusters.
```{r}
writeSpatialShape(gwr_df, 'gwr_shapefile')
```



# Load and prepare outlines of significant clusters for mapping
```{r}
sig_clusters <- readOGR(dsn = '.',
                   layer = 'GWR_dissolved_clusters',
                   verbose = T) %>% 
  tidy() %>% 
  filter(lat < 50 & lat > 20) %>% 
  filter(long > -130 & long < -60)
```


# Prepare files for ggplot by limiting the map to only the 48 contiguous states
```{r}
gwr_tidied <- gwr_tidied %>%
  filter(lat < 50 & lat > 20) %>% 
  filter(long > -130 & long < -60) %>% 
  mutate(gini_tv_bin = if_else(abs(gini_TV) > 1.96, T, F))

county_tidied <- county_tidied %>% 
  filter(lat < 50 & lat > 20) %>% 
  filter(long > -130 & long < -60) %>% 
  mutate(gini = NA) %>% 
  mutate(gini_tv_bin = NA)

nation_tidied <- tidy(nation_map) %>% 
  filter(lat < 50 & lat > 20) %>% 
  filter(long > -130 & long < -60)

map_long <- c(gwr_tidied$long, county_tidied$long)
map_lat <- c(gwr_tidied$lat, county_tidied$lat)
```



# Map the results

Maps were exported as images with dimensions 3300x2550
```{r}
# Beta coefficients
ggplot()  + 
  geom_polygon(data = gwr_tidied,
                        aes(x = long,
                            y = lat,
                            group = group,
                            fill = gwr_tidied$gini)) +
  geom_polygon(data = county_tidied,
               aes(x = long,
                   y = lat,
                   group = group)) +
  geom_polygon(data = sig_clusters,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = 'black',
               size = 4) +
  geom_polygon(data = nation_tidied,
                        aes(x = long,
                            y = lat,
                            group = group),
                            color = 'black',
                        fill = NA,
                        size = 1) +
  scale_fill_gradient2() +
  labs(fill = expression(beta),
       x = "Longitude",
       y = "Latitude") +
  theme_bw() +
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) +
  coord_map('albers', 
            lat0 = 30, lat1 = 40, 
            xlim = c(min(map_long) - 0.0001, max(map_long) + 0.0001),
            ylim = c(min(map_lat) - 0.0001, max(map_lat) + 0.0001)) +
  theme(legend.key.size = unit(3, "lines"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25),
        plot.margin = unit(c(.1, .1, .1, .1), 'cm'))
  

# T-values
ggplot() + geom_polygon(data = gwr_tidied,
                        aes(x = long,
                            y = lat,
                            group = group,
                            fill = gwr_tidied$gini_tv_bin)) +
  geom_polygon(data = county_tidied,
               aes(x = long,
                   y = lat,
                   group = group)) +
  geom_polygon(data = nation_tidied,
                        aes(x = long,
                            y = lat,
                            group = group),
                            color = 'black',
                        fill = NA,
                        size = 1) +
  labs(x = "Longitude",
       y = "Latitude") +
  scale_fill_discrete(na.value = "grey50",
                      breaks = c(T, F),
                      labels = c('Significant', 'Not significant'),
                      name = expression(paste('Statistical significance (',
                                   italic(p), ' < 0.05)'))) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) +
  coord_map('albers', 
            lat0 = 30, lat1 = 40, 
            xlim = c(min(map_long) - 0.0001, max(map_long) + 0.0001),
            ylim = c(min(map_lat) - 0.0001, max(map_lat) + 0.0001)) +
  theme(legend.key.size = unit(3, "lines"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25),
        plot.margin = unit(c(.1, .1, .1, .1), 'cm'))

```