---
title: "Mapping"
author: "Felix Tran"
date: "November 19, 2018"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(sp)
library(rgdal)
library(rgeos)
library(ggplot2)
library(broom)
library(RColorBrewer)
library(gridExtra)
library(mapproj)
```

This file contains code to generate maps of the distribution of the variables
used in my thesis. 

# Plot of counties with usable suicide rates

Read in the shp file for US counties and the thesis dataset
```{r}
county_map <- readOGR(dsn = 'cb_2016_us_county_500k',
                   layer = 'cb_2016_us_county_500k',
                   verbose = T)

final_df <- readr::read_csv('./Cleaned data/County/final_df.csv')
unreliable_rates <- which(final_df$unreliable == T)
final_df$crude_rate[unreliable_rates] <- NA

nation_map <- readOGR(dsn = 'cb_2016_us_state_500k',
                   layer = 'cb_2016_us_state_500k',
                   verbose = T)
```

1. Convert the FIPS code into numeric type to allow merging with final_df.

2. Merge map file with final_df

3. Grab ID's of each polygon in map file and retain them so we can merge
map data with the tidied map dataset

4. Remove observations beyond a certain longitude and latitude (not sure what 
these are but they are definitely not part of the 50 U.S. states). Also filter
out observations for Puerto Rico.

5. Plot
```{r}
county_map@data$GEOID <- as.numeric(as.character(county_map@data$GEOID))

county_map_merged <- merge(county_map, final_df, by.x = 'GEOID', by.y = 'county_code')

county_map_merged$id <- sapply(slot(county_map_merged,  "polygons"),
                           function(x) slot(x, "ID"))

county_map_tidied <- tidy(county_map_merged)

county_map_tidied <- merge(county_map_tidied, county_map_merged@data, by = "id",
                        all.x = T) %>% 
  filter(lat < 50 & lat > 20) %>% 
  filter(long > -130 & long < -60) 

nation_tidied <- tidy(nation_map) %>% 
  filter(lat < 50 & lat > 20) %>% 
  filter(long > -130 & long < -60) 
```



# Plot of counties with usable rates vs. unusable rates
```{r}
ggplot() +
  geom_polygon(data = county_map_tidied,
               aes(x = long,
                   y = lat,
                   group = group,
                   fill = usable),
               color = 'black') +
  scale_fill_discrete(c('red', 'blue')) +
  labs(title = 'USA counties and suicide rates',
       fill = 'Usable suicide rate') +
  theme_void()
```



# Plot of county-level suicide rates in the 48 contiguous states and DC

Map was exported with dimensions 3300x2550
```{r}
map_long <- county_map_tidied$long
map_lat <- county_map_tidied$lat

ggplot() + 
  geom_polygon(data = county_map_tidied,
               aes(x = long,
                   y = lat,
                   group = group,
                   fill = crude_rate)) +
  geom_polygon(data = nation_tidied,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               size = 0.5,
               color = 'black') +
  scale_fill_gradient(low = '#56B1F7', high = '#132B43') +
  labs(fill = "Suicide rate per\n 100,000 people",
       x = "Longitude",
       y = "Latitude") +
  theme_bw() +
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20)) +
  coord_map('albers', 
            lat0 = 30, lat1 = 40, 
            xlim = c(min(map_long) - 0.0001, max(map_long) + 0.0001),
            ylim = c(min(map_lat) - 0.0001, max(map_lat) + 0.0001)) +
  theme(legend.key.size = unit(3, "lines"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25),
        plot.margin = unit(c(.1, .1, .1, .1), 'cm'))
```



# Plot of county-level income inequality in the 48 contiguous states and DC
```{r}
ggplot() + 
  geom_polygon(data = county_map_tidied,
               aes(x = long,
                   y = lat,
                   group = group,
                   fill = gini)) +
  geom_polygon(data = nation_tidied,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               size = 0.1,
               color = 'black') +
  scale_fill_gradient(na.value = '#D3D3D3', low = '#56B1F7', high = '#132B43') +
  labs(fill = "Gini coefficient",
       x = "Longitude",
       y = "Latitude") +
  theme_bw() +
  theme(legend.title = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8)) +
  theme(legend.key.size = unit(0.75, "lines"),
        legend.text = element_text(size = 8)) +
  coord_map('albers', 
            lat0 = 30, lat1 = 40, 
            xlim = c(min(map_long) - 0.0001, max(map_long) + 0.0001),
            ylim = c(min(map_lat) - 0.0001, max(map_lat) + 0.0001))
```