---
title: "Mapping"
author: "Felix Tran"
date: "November 19, 2018"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(sp)
library(rgdal)
library(rgeos)
library(ggplot2)
library(broom)
library(RColorBrewer)
library(gridExtra)
```

This file contains code to generate maps of the distribution of the variables
used in my thesis. 

# Plot of counties with usable suicide rates

Read in the shp file for US counties and the thesis dataset
```{r}
usa_map <- readOGR(dsn = 'cb_2016_us_county_500k',
                   layer = 'cb_2016_us_county_500k',
                   verbose = T)

head(usa_map@data)

final_df <- readr::read_csv('./Cleaned data/final_df.csv')
```

1. Convert the FIPS code into numeric type to allow merging with final_df.

2. Merge map file with final_df

3. Grab ID's of each polygon in map file and retain them so we can merge
map data with the tidied map dataset

4. Remove observations beyond a certain longitude and latitude (not sure what 
these are but they are definitely not part of the 50 U.S. states). Also filter
out observations for Puerto Rico.

5. Plot
```{r}
usa_map@data$GEOID <- as.numeric(as.character(usa_map@data$GEOID))

usa_map_merged <- merge(usa_map, final_df, by.x = 'GEOID', by.y = 'county_code')

usa_map_merged$id <- sapply(slot(usa_map_merged,  "polygons"),
                           function(x) slot(x, "ID"))

usa_map_tidied <- tidy(usa_map_merged)

usa_map_tidied <- merge(usa_map_tidied, usa_map_merged@data, by = "id",
                        all.x = T)

usa_map_tidied <- usa_map_tidied %>% 
  filter(long < 0 & lat > 12) %>% 
  filter(!(lat < 20 & long > -75))
```

##### Plot of counties with usable rates vs. unusable rates
```{r}
ggplot() +
  geom_polygon(data = usa_map_tidied,
               aes(x = long,
                   y = lat,
                   group = group,
                   fill = usable),
               color = 'black') +
  scale_fill_discrete(c('red', 'blue')) +
  labs(title = 'USA counties and suicide rates',
       fill = 'Usable suicide rate') +
  theme_void()

# counties_plot
# ggsave('USA counties.jpg', counties_plot)
```

##### Plot showing distribution of age-adjusted suicide rates
```{r}
ggplot() +
  geom_polygon(data = usa_map_tidied,
               aes(x = long,
                   y = lat,
                   group = group,
                   fill = age_adjusted_rate),
               color = 'black') +
  labs(title = 'USA counties and suicide rates') 
```

##### Plot of distribution of age-adjusted suicide rates in the 48 contiguous
##### states and DC
```{r}
usa_map_tidied %>% 
  filter(!(state == "Alaska" | state == "Hawaii")) %>% 
  ggplot(aes(x = long, y = lat, group = group, fill = age_adjusted_rate)) +
  geom_polygon(color = 'black')
```

