---
title: "Multicollinearity"
author: "Felix Tran"
date: "December 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(data.table)
library(stats)
library(readr)
```

This file contains code to construct a multicollinearity matrix to assess 
collinearity among the covariates/exposure.

# Multicollinearity matrix - County

1. Read in the county dataset

2. Keep only the covariates to assess multicollinearity

3. Save result as a csv
```{r}
county_df <- read_csv('./Cleaned data/County/final_df.csv') %>% 
  filter(usable == T) %>% 
  select(gini,
         median_age, age_15_24_prop, age_25_44_prop, age_45_64_prop, geriatric_prop,
         white_prop, black_prop, native_prop, asian_prop, hispanic_prop,
         median_income, poverty_prop, unemployed_prop, blue_collar_prop,
         below_hs_prop, hs_prop, some_college_prop, college_prop
         )

county_multicol_matrix <- round(cor(county_df), digits = 3)
write.csv(county_multicol_matrix, file = './Tables/Multicollinearity table.csv')
```


# Multicollinearity matrix - State
1. Read in the county dataset

2. Keep only the covariates to assess multicollinearity

3. Urban-ruralness of a state can be calculated by the proportion of counties
with a particular NCHS urban-rural rode within a state. This calculated and 
merged wtih the state dataset

3. Save result as a csv
```{r}
urban_rural_df <- read_csv('./Cleaned data/County/final_df.csv') %>% 
  select(state, urban_rural_code) %>% 
  mutate(urban_rural_recoded = as.integer((urban_rural_code - 1) / 2)) %>% 
  group_by(state, urban_rural_recoded) %>% 
  summarize(counts = n()) %>% 
  group_by(state) %>% 
  mutate(county_total = sum(counts)) %>% 
  ungroup() %>% 
  mutate(proportion = round(counts / county_total, digits = 3)) %>% 
  select(-counts, -county_total) %>% 
  spread(key = urban_rural_recoded, value = proportion) %>% 
  rename(rural_prop = '2',
         urban_prop = '0',
         in_between_prop = '1') %>% 
  mutate(rural_prop = if_else(is.na(rural_prop), 0, rural_prop),
         in_between_prop = if_else(is.na(in_between_prop), 0, in_between_prop),
         urban_prop = if_else(is.na(urban_prop), 0, urban_prop))
```

```{r}
state_df <- read_csv('./Cleaned data/State/state_final_df.csv') %>% 
  select(gini,
         median_age, age_15_24_prop, age_25_44_prop, age_45_64_prop, geriatric_prop,
         white_prop, black_prop, native_prop, asian_prop, hispanic_prop,
         median_income, poverty_prop, unemployed_prop, blue_collar_prop,
         below_hs_prop, hs_prop, some_college_prop, college_prop, 
         ffl_per_100000, 
         state
         )

state_df <- left_join(state_df, urban_rural_df, by = 'state') %>% 
  select(-state)
```

```{r}
state_multicol_matrix <- round(cor(state_df), digits = 3)
write.csv(state_multicol_matrix, file = './Tables/STATE_multicollinearity table.csv')
```



# Checking for potential differences in multicollinearity between levels

Taking the variables the 2 levels have in common, we can examine if the 
correlations between variables drastically differ by level. The level 
(county or state) does not seem to drastically change the correlations between
variables.
```{r}
diff_matrix <- as_data_frame(county_multicol_matrix) - 
  as_data_frame(state_multicol_matrix[1:19, 1:19])
```

