---
title: "STATE_data_cleaning"
author: "Felix Tran"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(readr)
library(janitor)
library(data.table)
```

This file contains code for cleaning and merging raw state-level data into a 
final dataset.

# Main exposure - Income inequality

Read in file on exposure data and rename variables. Removed Puerto Rico from
analyses because its history as a territory make it difficult to compare to
states
```{r}
gini_df <- read_csv('./Raw data/State/ACS 2012-2016 State gini.csv', skip = 1) %>% 
  clean_names() %>% 
  select(-id, -margin_of_error_gini_index) %>% 
  rename(state_code = id2,
         state = geography,
         gini = estimate_gini_index) %>% 
  filter(state != 'Puerto Rico') %>% 
  mutate(state_code = as.integer(state_code))
```


# Main outcome - Suicide rates

Read in file on outcome data
```{r}
suicide_df <- read_tsv('./Raw data/State/State Compressed Mortality, 1999-2016.txt') %>% 
  clean_names() %>% 
  filter(!is.na(state)) %>%
  select(state_code, crude_rate, age_adjusted_rate) %>% 
  mutate(state_code = as.integer(state_code))
```


# Covariates


### Demographics - Race/ethnicity, gender, and race
```{r}
demographic_df <- read_csv('./Raw data/State/ACS 2012-2016 State demographics.csv') %>% 
  clean_names() %>% 
  filter(geo_display_label != 'Puerto Rico') %>% 
  select(state_code = geo_id2,
         total_pop = hc01_vc03,
         male_prop = hc03_vc04,
         white_prop = hc03_vc94,
         black_prop = hc03_vc95,
         native_prop = hc03_vc96,
         asian_prop = hc03_vc97,
         hispanic_prop = hc03_vc88,
         median_age = hc01_vc23,
         geriatric_prop = hc03_vc29,
         hc03_vc11, hc03_vc12,
         hc03_vc13, hc03_vc14, 
         hc03_vc15, hc03_vc16, hc03_vc17) %>% 
  mutate(state_code = as.integer(state_code),
         age_15_24_prop = as.numeric(hc03_vc11) + as.numeric(hc03_vc12),
         age_25_44_prop = as.numeric(hc03_vc13) + as.numeric(hc03_vc14),
         age_45_64_prop = as.numeric(hc03_vc15) + as.numeric(hc03_vc16) +
           as.numeric(hc03_vc17)) %>% 
  select(-c(hc03_vc11, hc03_vc12, hc03_vc13, hc03_vc14, hc03_vc15,
            hc03_vc16, hc03_vc17))
demographic_df <- demographic_df[-1, ]
```


### Educational attainment
```{r}
edu_df <- read_csv('./Raw data/State/ACS 2012-2016 State education.csv') %>% 
  clean_names() %>% 
  filter(geo_display_label != 'Puerto Rico') %>% 
  select(state_code = geo_id2,
         hc02_est_vc09,
         hc02_est_vc10,
         hs_prop = hc02_est_vc11,
         hc02_est_vc12,
         hc02_est_vc13,
         college_prop = hc02_est_vc18) %>% 
  mutate(state_code = as.integer(state_code),
         below_hs_prop = as.numeric(hc02_est_vc09) + as.numeric(hc02_est_vc10),
         some_college_prop = as.numeric(hc02_est_vc12) + as.numeric(hc02_est_vc13)) %>% 
  select(-c(hc02_est_vc09, hc02_est_vc10, hc02_est_vc12, hc02_est_vc13))
edu_df <- edu_df[-1, ]
```


### Economic characteristics - Median income, poverty, unemployment, occupation
```{r}
econ_df <- read_csv('./Raw data/State/Acs 2012-2016 State economics.csv') %>% 
  clean_names() %>% 
  filter(geo_display_label != 'Puerto Rico') %>% 
  select(state_code = geo_id2,
         median_income = hc01_vc86,
         poverty_prop = hc03_vc171,
         unemployed_prop = hc03_vc12,
         hc03_vc44, 
         hc03_vc45) %>% 
  mutate(state_code = as.integer(state_code),
         blue_collar_prop = as.numeric(hc03_vc44) + as.numeric(hc03_vc45)) %>% 
  select(-hc03_vc44, -hc03_vc45)
econ_df <- econ_df[-1, ]
```


### Federally licensed firearms dealers per capita (proxy for gun ownership)

As a proxy for gun ownership, I am estimating the number of gun dealers per
capita in each state. I downloaded a pdf file listing the total number of 
federal firearm licensees (ffl) in each state for the month of July in 2014. 
Because ATF did not have data on FFL's before 2014, I took the middle of the 
study timeperiod (2012-2016) as an estimate for the total period.

After reading in the table, I summed up the number of type 1 and 2 FFL's.

Type 1 = Dealer in Firearms Other Than Destructive Devices (Includes Gunsmiths)
Type 2 = Pawnbroker in Firearms Other Than Destructive Devices

I then converted state abbreviations into their full names for merging with
other data, then filtered out entities (e.g. Puerto Rico, American Samoa) that
are not states.
```{r}
raw_gun_data <- tabulizer::extract_tables('./Raw data/State/ffl-type-by-state-july-2014.pdf',
                                          output = 'data.frame')

gun_df <- as_tibble(rbind(raw_gun_data[[1]], raw_gun_data[[2]])) %>% 
  clean_names() %>% 
  mutate(dealer = as.numeric(gsub(",", "", x01)),
         pawnshop = as.numeric(gsub(",", "", x02)),
         ffl = dealer + pawnshop) %>% 
  select(state, ffl)

gun_df$state <- openintro::abbr2state(gun_df$state)
gun_df <- gun_df %>% 
  filter(!is.na(state))
```



# Combining datasets

I  divided the sum of type 1 and 2 FFl's by the state population and 
multipled the quotient by 10,000 to estimate number of gunshops per 100,000
residents in a given state.
```{r merge datasets}
state_final_df <- left_join(gini_df, suicide_df, by = "state_code") %>% 
  left_join(., gun_df, by = "state") %>% 
  left_join(., econ_df, by = "state_code") %>% 
  left_join(., edu_df, by = "state_code") %>% 
  left_join(., demographic_df, by = "state_code") %>% 
  mutate(ffl_per_100000 = round((ffl / as.numeric(total_pop)) * 100000, 
                               digits = 2))
```



# Saving final merged dataset
```{r}
write_csv(state_final_df, path = "./Cleaned data/State/state_final_df.csv")
```

